---
title: "p8105_hw2_wz2507"
author: "Wurongyan Zhang"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
```{r, message=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
```


```{r}
# create the data set
trash <- read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408") %>% 
  janitor::clean_names() %>% #clean names
  drop_na(dumpster) %>% #drop non dumpster data
  mutate(
    sports_balls = as.integer( sports_balls)) 
#round the number to integers
  
  
```

```{r}
# create precipatation data for 2018
trash18 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 3, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(Year = "2018") %>% #change the year to 2018
  mutate(month = month.name) %>% #change to full month names
  drop_na()  #omit rows without data

```

```{r}
# create precipatation data for 2018
trash17 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14")%>% 
  janitor::clean_names() %>% 
  mutate(month = month.name) %>% #change to full month names
  mutate(Year = "2017") #change the year to 2018
  
```

```{r, warning=FALSE}
#combine precipatation data for 2017 and 2018
precip1718 <- as.tibble(rbind(trash17, trash18))

```




```{r}
ball17 <- tibble(x = pull(trash, year ), y = pull(trash, sports_balls )) %>% 
  filter(x == "2017")
```

Comment: The number of observations in Mr.Trash Wheel data set is `r nrow(trash)` with 14 variables and in precipatation data set is `r nrow(precip1718)` with 3 variables. 
key variables??
The total precipatation in 2018 is `r sum(pull(trash18, total))`. The median of sports ball is `r median(pull(ball17, y))`.
The key variables for trash includes dumpter number(dumpster), data of collection, amount of total litter(weight_tons, volume_cubic_yards) and litter type(plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_balls, etc.). For the precipatation data, it includes month, year and the total precipitation amounts.




## Problem 2

```{r, message=FALSE}
# create pols data set
pols <- read_csv("data/pols-month.csv") %>% 
  #convert the date into year, month and day
  separate(mon, into = c("year", "month", "day")) %>% 
  #change month into full name
  mutate(month = as.integer(month)) %>% 
  mutate(month = month.name[month]) %>% 
  #specify if the presidwent is democrat or republican
  mutate(prez_dem = recode(prez_dem, "1" = "democrat", "0" = "republican")) %>% 
  mutate(president =  prez_dem) %>% 
  #delete prez_dem, prez_gop and day
  select(-prez_dem, -prez_gop) %>% 
  select(-day)%>% 
  mutate(year = as.integer(year))
```

```{r, message=FALSE}
# create snp data set
snp <- read_csv("data/snp.csv") %>% 
  #convert the date into year, month and day
  separate(date, into = c("month", "day", "year")) %>%
  #change month into full name
  mutate(month = as.integer(month)) %>%
  #arrange the order by year, month and others
  arrange(year, month) %>% 
  select(year, month, everything()) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(year = as.integer(year))

```

```{r, message=FALSE}
#create unemployment data set
une <- read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  # change month to a column
  pivot_longer(jan:dec,
               names_to = "month", 
               values_to = "unemployment rate") %>%
  #change month into full name
  mutate(month = factor(month, labels = month.name))%>%
  #arrange the data by year, month and others
  arrange(year, month) %>% 
  select(year, month, everything())

```


```{r, message=FALSE}
#combine those three data sets
join = full_join(pols, une)
all = full_join(join, snp)
```


Comments: The dimension of the data set after combination is `r nrow(all)` entries with `r ncol(all)` total columns(if we delete prez_gop, prez_dem and day). The range of years is 1947 to 2015. However, pols data starts from 1947, snp data starts from 1950 and une data starts from 1947. Therefore, the combined data range from 1947 to 2015. The names of key variables are --

for pols data:

year: year of the count

month: month of the count of the year

rep_dem: the number of democratic representatives on the associated date

gov_gop: the number of republican governors on the associated date

sen_gop: the number of republican senators on the associated date

rep_gop: the number of republican representatives on the associated date

gov_dem: the number of democratic governors on the associated date

sen_dem: the number of democratic senators on the associated date

president:  indicator of whether the president was democratic or republican

for snp data:

year: the date of the observation

month: the month of the observation

day: the day of the observation

close: the closing values of the S&P stock index on the associated date

for unemployment data:

year: the year of the measurements on that row

month: the month of the measurements on that row

rate: unemployment rate

## Problem 3

```{r}
# create the function of capital letter of the first names
first <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2), sep = "")
}
```
  
```{r, message=FALSE}
# create the data set
# change column names
baby <- read_csv("data/Popular_Baby_Names.csv", skip = 1,col_names =c("birth_year", "gender", "race", "first_name", "count","rank")) %>% 
  # change the first name into capital letter of the first character with rest of lower case
  mutate(
    first_name = str_to_lower(first_name), first_name = first(first_name)) %>%
  # lower the case for gender
  # unify race names
  mutate(
    gender = str_to_lower(gender), race = str_to_lower(race)) %>% 
  mutate(race = replace(race, race == "asian and paci", "asian and pacific islander"), 
         race = replace(race,race == "black non hisp","black non hispanic"), 
         race = replace(race,race == "white non hisp","white non hispanic")) %>% 
  #delete duplicate rows
  distinct()
         
  
```
# i.
  
```{r}
# rank of Olivia
# create the data set, pull out data for Olivia
olivia = baby[pull(baby, first_name) == "Olivia",] %>% 
  # arrange the set by birth year
  arrange(birth_year)
```

```{r}
# create the table
pivot_wider(
  olivia, id_cols = "birth_year", names_from = "race", values_from = "rank"
)
```

# ii.
```{r}
# rank the male children 
male_name = baby[pull(baby, gender) == "male",] %>% 
  group_by(birth_year, race) %>% 
  slice(which.min(rank))
# create the table
pivot_wider(
  male_name, id_cols = "birth_year", names_from = "race", values_from = "first_name"
)
```

# iii.
```{r}
# pull the data of male, white born in 2016
male = baby[pull(baby, gender) == "male",] 

white = male[pull(male, race) == "white non hispanic",] 
white16 = white[pull(white, birth_year) == "2016",]
```

```{r}
# create the ggplot

white16 %>% 
  ggplot(aes(x = rank, y = count )) + geom_point() 
```

































