---
title: "p8105_hw2_wz2507"
author: "Wurongyan Zhang"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
```{r}
library(readxl)
library(tidyverse)
```


```{r}
trash <- read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(
    sports_balls = as.integer( sports_balls))
  
  
```

```{r}
trash18 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 3, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(Year = "2018") %>%
  mutate(month = month.name) %>% 
  drop_na()  

```

```{r}
trash17 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14")%>% 
  janitor::clean_names() %>% 
  mutate(month = month.name) %>% 
  mutate(Year = "2017") 
  
```

```{r}
precip1718 <- as.tibble(rbind(trash17, trash18))

```


Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017.

```{r}
ball17 <- tibble(x = pull(trash, year ), y = pull(trash, sports_balls )) %>% 
  filter(x == "2017")
```

The number of observations in Mr.Trash Wheel data set is 344 with 14 variables and in precipatation data set is 19 with 3 variables. 
key variables??
The total precipatation in 2018 is `r sum(pull(trash18, total))`. The median of sports ball is `r median(pull(ball17, y))`.


## Problem 2

```{r}
pols <- read_csv("data/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = as.integer(month)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(prez_dem = recode(prez_dem, "1" = "democrat", "0" = "republican")) %>% 
  mutate(president =  prez_dem) %>% 
  select(-prez_dem, -prez_gop) %>% 
  select(-day)%>% 
  mutate(year = as.integer(year))
```

```{r}
snp <- read_csv("data/snp.csv") %>% 
  separate(date, into = c("month", "day", "year")) %>%
  mutate(month = as.integer(month)) %>%
  arrange(year, month) %>% 
  select(year, month, everything()) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(year = as.integer(year))

```

```{r}
une <- read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec,
               names_to = "month", 
               values_to = "unemployment rate") %>%
  mutate(month = factor(month, labels = month.name))%>%
  arrange(year, month) %>% 
  select(year, month, everything())

```


```{r}
join = full_join(pols, une)
all = full_join(join, snp)
```
Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

The dimension of the data set is 828 entries with 12?? total columns(if we delete prez_gop, prez_dem and day). The range of years is 1947 to 2015. However, pols data starts from 1947, snp data starts from 1950 and une data starts from 1947. The names of key variables are --

for pols data: 
mon: date of the count
rep_dem: the number of democratic representatives on the associated date
gov_gop: the number of republican governors on the associated date
sen_gop: the number of republican senators on the associated date
rep_gop: the number of republican representatives on the associated date
gov_dem: the number of democratic governors on the associated date
sen_dem: the number of democratic senators on the associated date
president:  indicator of whether the president was democratic or republican

for snp data:
year: the date of the observation
month: the month of the observation
day: the day of the observation
close: the closing values of the S&P stock index on the associated date

For unemployment data:
year: the year of the measurements on that row
month: the month of the measurements on that row
each month: percentage of unemployment in each month of the associated year
rate: unemployment rate

## Problem 3

```{r}
first <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2), sep = "")
}
```
  
```{r}
baby <- read_csv("data/Popular_Baby_Names.csv", skip = 1,col_names =c("birth_year", "gender", "race", "first_name", "count","rank")) %>% 
  mutate(
    first_name = str_to_lower(first_name), first_name = first(first_name)) %>%
  mutate(
    gender = str_to_lower(gender), race = str_to_lower(race)) %>% 
  mutate(race = replace(race, race == "asian and paci", "asian and pacific islander"), 
         race = replace(race,race == "black non hisp","black non hispanic"), 
         race = replace(race,race == "white non hisp","white non hispanic")) %>% 
  distinct()
         
  
```
most_popular_male_names <- names[pull(names,gender)=='male',] %>% 
  group_by(birth_yr,race) %>% 
  slice(which.min(rank))
  
```{r}
olivia = baby[pull(baby, first_name) == "Olivia",] %>% 
  arrange(birth_year)
```

```{r}
pivot_wider(
  olivia, id_cols = "birth_year", names_from = "race", values_from = "rank"
)
```

```{r}
male_name = baby[pull(baby, gender) == "male",] %>% 
  group_by(birth_year, race) %>% 
  slice(which.min(rank))

pivot_wider(
  male_name, id_cols = "birth_year", names_from = "race", values_from = "first_name"
)
```

```{r}
male = baby[pull(baby, gender) == "male",] 

white = male[pull(male, race) == "white non hispanic",] 
white16 = white[pull(white, birth_year) == "2016",]
```

```{r}
library(ggplot2)
white16 %>% 
  ggplot(aes(x = rank, y = count )) + geom_point() 
```

































