---
title: "p8105_hw2_wz2507"
author: "Wurongyan Zhang"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
```{r, message=FALSE}
# load library
library(readxl)
library(tidyverse)
library(ggplot2)
```


```{r}
# Import the trash wheel data set
trash <- read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408") %>% 
  janitor::clean_names() %>% #clean names
  drop_na(dumpster) %>% #drop non dumpster data
  #round the number of sports balls
  mutate(
    sports_balls = as.integer( sports_balls)) 

  
  
```

```{r}
# import precipatation data for 2018
trash18 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 3, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(Year = "2018") %>% #add the variable year
  mutate(month = month.name) %>% #change to full month names
  drop_na()  #omit rows without data

```

```{r}
# create precipatation data for 2018
trash17 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14")%>% 
  janitor::clean_names() %>% 
  mutate(month = month.name) %>% #change to full month names
  mutate(Year = "2017") #add the variable year
  
```

```{r, warning=FALSE}
#combine precipatation data for 2017 and 2018
precip1718 <- as.tibble(rbind(trash17, trash18))

```

```{r}
# filter the data set only contains sports balls in 2017 in order to find the median
ball17 <- tibble(x = pull(trash, year ), y = pull(trash, sports_balls )) %>% 
  filter(x == "2017")
```

Comment: The number of observations in Mr.Trash Wheel data set is `r nrow(trash)` with `r ncol(trash)` variables and the number of observations in precipatation data set is `r nrow(precip1718)` with `r ncol(precip1718)` variables. 

The total precipatation in 2018 is `r sum(pull(trash18, total))`. The median of sports ball is `r median(pull(ball17, y))`.
The key variables for trash data set include dumpter number (dumpster), date of collection, amount of total litter (weight_tons, volume_cubic_yards) and litter types (plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_balls, etc.). For the precipatation data, it includes month, year and the total precipitation amounts. The name of the columns are shown below:
```{r}
head(trash)
head(precip1718)
```





## Problem 2

```{r, message=FALSE}
# import pols data set
pols <- read_csv("data/pols-month.csv") %>% 
  #separate the date into year, month and day to 3 variables
  separate(mon, into = c("year", "month", "day")) %>% 
  #change month into full name
  mutate(month = as.integer(month)) %>% 
  mutate(month = month.name[month]) %>% 
  #specify if the president is democrat or republican by adding a new variable president
  mutate(prez_dem = recode(prez_dem, "1" = "democrat", "0" = "republican")) %>% 
  mutate(president =  prez_dem) %>% 
  #delete prez_dem, prez_gop and day
  select(-prez_dem, -prez_gop) %>% 
  select(-day)%>% 
  mutate(year = as.integer(year))
```

```{r, message=FALSE}
# import snp data set
snp <- read_csv("data/snp.csv") %>% 
  #convert the date into year, month and day
  separate(date, into = c("month", "day", "year")) %>%
  #change month into full name
  mutate(month = as.integer(month)) %>%
  #arrange the order by year, month and others
  arrange(year, month) %>% 
  select(year, month, everything()) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(year = as.integer(year))

```

```{r, message=FALSE}
# import unemployment data set
une <- read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  # change month to a column
  pivot_longer(jan:dec,
               names_to = "month", 
               values_to = "unemployment rate") %>%
  #change month into full name
  mutate(month = factor(month, labels = month.name))%>%
  #arrange the data by year, month and others
  arrange(year, month) %>% 
  select(year, month, everything())

```


```{r, message=FALSE, warning=FALSE}
#combine those three data sets
join = full_join(pols, une)
all = full_join(join, snp)
```


Comments: The dimension of the data set after combination is `r nrow(all)` entries with `r ncol(all)` total columns (if we delete prez_gop, prez_dem and day). The range of years is from 1947 to 2015. Specifically for each data set, the dimension of pols data set is `r nrow(pols)` entries with `r ncol(pols)` columns ranging from 1947 to 2015; the dimension of snp data is `r nrow(snp)` entries with `r ncol(snp)` columns ranging from 1950 to 2015 ; the dimension of unemplyment data is `r nrow(une)` entries with `r ncol(une)` columns ranging from 1948 to 2015. The names of key variables are --


year: year of the count; month: month of the count of the year; day: the day of the observation

rep_dem: the number of democratic representatives on the associated date

gov_gop: the number of republican governors on the associated date

sen_gop: the number of republican senators on the associated date

rep_gop: the number of republican representatives on the associated date

gov_dem: the number of democratic governors on the associated date

sen_dem: the number of democratic senators on the associated date

president:  indicator of whether the president was democratic or republican

close: the closing values of the S&P stock index on the associated date

rate: unemployment rate

## Problem 3

```{r}
# create the function of capital letter of the first names
first <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2), sep = "")
}
```
  
```{r, message=FALSE}
# import the baby name data set
# change column names
baby <- read_csv("data/Popular_Baby_Names.csv", skip = 1,col_names =c("birth_year", "gender", "race", "first_name", "count","rank")) %>% 
  # capitalize the first character of the first names 
  mutate(
    first_name = str_to_lower(first_name), first_name = first(first_name)) %>%
  # lower the case for gender
  # unify race names
  mutate(
    gender = str_to_lower(gender), race = str_to_lower(race)) %>% 
  mutate(race = replace(race, race == "asian and paci", "asian and pacific islander"), 
         race = replace(race,race == "black non hisp","black non hispanic"), 
         race = replace(race,race == "white non hisp","white non hispanic")) %>% 
  #delete duplicate rows
  distinct()
         
  
```
# i.
  
```{r}
# rank of Olivia
# create the data set, pull out data for Olivia
olivia = baby[pull(baby, first_name) == "Olivia",] %>% 
  # arrange the set by birth year
  arrange(birth_year)
```

```{r}
# create the table to show the rank of Olivia
pivot_wider(olivia, id_cols = "race", names_from = "birth_year", values_from = "rank") 
```


# ii.
```{r}
# rank the male children 
male_name = baby[pull(baby, gender) == "male",] %>% 
  group_by(birth_year, race) %>% 
  slice(which.min(rank))
# create the table
pivot_wider(male_name, id_cols = "race",names_from = "birth_year",values_from = "first_name")
```



# iii.
```{r}
# pull the data of male, white born in 2016
male = baby[pull(baby, gender) == "male",] 

white = male[pull(male, race) == "white non hispanic",] 
white16 = white[pull(white, birth_year) == "2016",]
```

```{r}
# create the ggplot showing the rank

plot = white16 %>% 
  ggplot(aes(x = rank, y = count )) + geom_point() +labs(title = "number of children with a name against the rank in popularity of that name", x ="rank", y = "number of children")
plot
ggsave("plot.pdf")

```

































